```
*   reasons for saving: 1=cant save, 2=education, 3=family, 4=home,
    5=purchases, 6=retirement, 7=liquidity/the future, 8=investment,
    9=no particular reason;
*   NOTE: multiple saving reasons may be reported: here choosing only
    first (most important) reason;
    ARRAY SAVRES {*} SAVRES1-SAVRES9;
    ARRAY RES {*} X3006 /*X3007*/ ;
    DO I=1 TO DIM(SAVRES);
      SAVRES{I}=0;
    END;
    DO I=1 TO DIM(RES);
      IF (RES{I} IN (-2 -1)) THEN SAVRES{1}=1;
      ELSE IF (RES{I} IN (1 2)) THEN SAVRES{2}=1;
      ELSE IF (RES{I} IN (3 5 6)) THEN SAVRES{3}=1;
      ELSE IF (RES{I} EQ 11) THEN SAVRES{4}=1;
      ELSE IF (RES{I} IN (12 13 14 15 16 27 29 30 9 18 20 41)) THEN
        SAVRES{5}=1;
      ELSE IF (RES{I} IN (17 22)) THEN SAVRES{6}=1;
      ELSE IF (RES{I} IN (23 24 25 32 92 93)) THEN SAVRES{7}=1;
      ELSE IF (RES{I} IN (21 26 28)) THEN SAVRES{8}=1;
      ELSE IF (RES{I} IN (31 33 40 90 91 -7)) THEN SAVRES{9}=1;
      ELSE PUT "UNCLASSIFIED SAVING REASON! " &ID= RES{I}=;
    END;
/*
    1.  Children's education; education of grandchildren
    2.  Own education; spouse's education; education -- NA for whom
    3.  "For the children/family"  -- NFS; "to help the kids
        out"; estate
    5.  Wedding, Bar Mitzvah, and other ceremonies (except 17)
    6.  To have children/a family
    9.  To move (except 11)
   11.  Buying own house (code summer cottage in 12)
   12.  Purchase of cottage or second home for own use
   13.  Buy a car, boat or other vehicle
   14.  Home improvements/repairs
   15.  To travel; take vacations; take other time off
   16.  Buy durable household goods, appliances, home
        furnishings; hobby and recreational items; for other
        purchases not codable above or not further specified;
        "buy things when we need/want them"; moving/special occasions
   17.  Burial/funeral expenses
   18.  Charitable or religious contributions
   20.  "To enjoy life"
   21.  Buying (investing in) own business/farm; equipment for
        business/farm
   22.  Retirement/old age
   23.  Reserves in case of unemployment
   24.  In case of illness; medical/dental expenses
   25.  Emergencies; "rainy days"; other unexpected needs; for
       "security" and independence
   26.  Investments reasons (to get interest, to be
        diversified, to buy other forms of assets)
   27.  To meet contractual commitments (debt repayment,
        insurance, taxes, etc.), to pay off house
   28.  "To get ahead;" to advance standard of living
   29.  Ordinary living expenses/bills
   31.  No reason (except 90, 91, 92)
   32.  "For the future"
   90.  Had extra income; saved becaused had the money left
        over -- no other purpose specified
   91.  Wise/prudent thing to do; good discipline to save; habit
   92.  Liquidity; to have cash available/on hand
   -1.  Don't/can't save; "have no money"
  ```
